### 2020-12-10 Thesis-Abstract.tex
Editorial improvements

Public debate has resumed over whether encryption systems should support alternative means of decryption intended for
law enforcement use, called exceptional access (EA). Rather than a renege on a solemn promise made at the end of the
1990s ``crypto war,'' this represents a valid reassessment of optimal policy in light of changing circumstances.
Achieving proper balance between privacy and access in the context of constantly changing society and technology is a
wicked problem that has and will evade a permanent solution. As policymakers consider next steps, it behooves the
technical community to stay engaged. Although the introduction of EA would inevitably introduce risk, the quality of the
technical and regulatory approach can make a substantial difference. Furthermore, if one considers ham-fisted
legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall
risk.

The root of the conflict lies in cryptography's dual role as enabler of radical privacy and cornerstone of security. The
emergence of strong encryption triggered the first crypto war, and its proliferation is causing the second. In the face
of both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked problems and promote
an iterative approach to the problem of encryption and EA. Along the way I illustrate the components of the debate in
argument maps and demonstrate the security risks with data flow diagrams, making clear distinctions for the separate
problems of EA for data in motion and data at rest.

>>>>

Public debate has resumed on the topic of exceptional access (EA), which is alternative means of decryption intended for
law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of the 1990s
``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing circumstances. The
imbalance between privacy, access, and security in the context of constantly changing society and technology is a wicked
problem that has and will continue to evade a permanent solution. As policymakers consider next steps, it is necessary
that the technical community to remain engaged. Although the introduction of EA would inevitably add risk, the quality
of the technical and regulatory approach can make a substantial difference. Furthermore, if one considers hard-line
legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce risk
overall.

The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
In response to both polarized and conciliatory voices, this paper analyzes strategies for confronting wicked problems
and proposes an iterative approach to the case of encryption and EA. Along the way, it illustrates the components of the
debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing on one
EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''

### 2020-12-10 Thesis-Chapter1-Introduction.tex
Editorial improvements

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
attempted solution has permanent and often unintended consequences, and is likely to exist in a pattern of chronic
policy failure.

@@

The conflict at the heart of the encryption and \ac{EA} debate is that the same cryptographic and design principles that
underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
communications and storage provides clear benefits and should be welcomed by all. In a world where information security
is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing
\cite{cox_2020} \cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk
should not be ignored.

The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}. Human rights
groups and technical leaders counter that a weakened encryption environment would fatally compromise privacy and
security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate continues,
sides have become entrenched, and both sides have engaged in disingenuous attacks against their opponents.

In the long term, increased regulation of the tech industry in general is inevitable, and regulation of encryption is
possible. Despite strong encryption's benefits, regulatory interest in the subject has not subdued. The form that
regulation will take in the field of encryption depends on the good faith efforts made to equitably balance all the
benefits and risks involved in deploying a cryptographic system. If the technical community acts, it can lead and shape
legislation instead of being subjected to it.

Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
community fails to act but regulators move forward, everyone may become subject to technically misguided laws that do
great harm. Bad \ac{EA} policy is just as much a threat as poor passwords. Due to this threat, as long as lawmakers
continue pursuing \ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and
respond to the arguments being presented.

@@

This paper accepts and builds on the following premises. These premises are not principles for potential \ac{EA}
designs, but the starting place for discussing the \ac{EA} debate.

@@

\principle{Cybersecurity is critical}{With current levels of reliance on computer information systems, cybersecurity is
critical to the wellbeing of society. Two important elements of cybersecurity are cryptography and system architecture.
Policy which supports security does not undercut cryptographic integrity or require high-risk architectures.}

\principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
privacy in all contexts is not one of them. While individuals under a limited government are entitled to an expectation
of privacy, an absolute right to privacy does not apply in all circumstances. Investigators should have access to some
classes of data. In particular, access to certain classes of digital data is important today and will be increasingly
important in the future.}

\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to \ac{masssurv},
potential for abuse, international consequences, economic impact, and need for transparency.}

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
right balance of risk given the threats under consideration---including the threat of regulatory action.}

@@

This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.--focused history of encryption regulation, technology, and conflict, I analyze strategies for tacking
\acp{wicked-prob} and introduce \acp{argmap} and threat modeling with data flow diagrams. In the following chapters, I
map the \acl{EA} arguments in detail and demonstrate the security risks with data flow diagrams and threat analysis,
focusing on one particular EA proposal. The thesis concludes with research and policy paths forward in light of valid
arguments and threats.

>>>>

Certain policy issues can be described as ``\acp{wicked-prob}.'' Originally coming from the field of design theory, the
term ``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, and measurements. Each
attempted solution has permanent and often unintended consequences, and is likely to exist in a pattern of chronic
policy failure.

@@

The conflict at the heart of the encryption and \ac{EA} debate is this: the same cryptographic and design principles
that underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
communications and storage provides clear benefits. In a world where information security is frightfully poor yet
increasingly important, the necessity of strong encryption cannot be understated. However, the privacy afforded by
certain encryption technologies hampers law enforcement investigations and hides wrongdoing \cite{cox_2020}
\cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk should not be
ignored.

The conversation regarding encryption has reached a stalemate. Governments and law enforcement agencies frequently cite
the need to access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}.
Human rights groups and technical leaders counter that a weakened encryption environment would fatally compromise
privacy and security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate
continues, many on both sides have become entrenched in their positions and have engaged in disingenuous attacks against
their opponents.

In the long term, increased regulation of the tech industry is inevitable and regulation of encryption is possible.
Despite the benefits of strong encryption, regulatory interest in the subject has not subdued. The form that regulation
will take in the field of encryption depends on the good faith efforts made to equitably balance the benefits and risks
involved in deploying a cryptographic system. If the technical community acts, it can lead and shape legislation rather
than be subjected to it.

Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
community fails to act but regulators move forward, everyone may become subject to technically misguided, harmful laws.
Bad policy is just as much a threat as weak passwords. Due to this threat, as long as lawmakers continue pursuing
\ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and respond to the
presented arguments.

@@

This paper accepts and builds on the following premises. These premises are not principles for potential \ac{EA}
designs, but the foundation for discussing the \ac{EA} debate.

@@

\principle{Cybersecurity is critical}{Due to modern culture's reliance on computer information systems, cybersecurity is
critical to the wellbeing of society. Two important elements of cybersecurity are cryptography and system architecture.
Policy that supports security does not undercut cryptographic integrity or require high-risk architectures.}

\principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
privacy in all contexts does not. While individuals under a limited government are entitled to an expectation of
privacy, an absolute right to privacy does not apply in all circumstances. Investigators should have access to some
classes of data. In particular, access to certain classes of digital data is important today and will become
increasingly important in the future.}

\principle{\Ac{EA} is an inherently complex problem}{The factors that make \ac{EA} a \ac{wicked-prob} require proposed
solutions to be analyzed at many levels. These factors include \ac{EA}'s relation to \ac{masssurv}, potential for abuse,
international consequences, economic impact, and need for transparency.}

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions; they do not even have
verifiably optimal solutions. Therefore, we cannot use perfection as the standard. To quote a famous security axiom,
insecurity cannot be destroyed, only moved around \cite{nrc_schneider_1999}. The \Ac{EA} problem, like all security
problems, involves finding the right balance of risk given the threats under consideration---including the threat of
regulatory action.}

@@

This thesis aims to clarify the arguments of the debate and the nature of the technical problem. After reproducing a
brief U.S.--focused history of encryption regulation, technology, and conflict, I analyze strategies for tackling
\acp{wicked-prob} and introduce \acp{argmap} and threat modeling with data flow diagrams. In the following chapters, I
map the \acl{EA} arguments in detail and demonstrate the security risks with data flow diagrams and threat analysis,
focusing on one EA proposal in particular. The thesis concludes with paths forward for research and policy that take the
arguments and threats discussed into consideration.
